\section{Reviewer 1}

\subsection{Major Comments}

\begin{enumerate}
	\item  The dataset suffers from two notable issues with regard to a fair comparison between ERGM, LSM, and the proposed AME approach. 
	\begin{itemize}
		\item The first is that each node’s latent position is likely to be correlated with the model terms of interest – a point raised by Cranmer et al. as well (pp. 248-9). In this case, key government players have lower Euclidean distances because they tend to be positioned in the center of the latent space. If this were not the case, would we see such a large divergence between the LSM and AME results? For example, would the AME perform much ``better'' than the LSM if the network data exhibited a less hierarchial/more egalitarian structure? 
			\begin{itemize}
				\item \textcolor{blue}{ \emph{
				To answer this question we conduct a simulation exercise that we have included in the appendix. Specifically, see page .
				}}
			\end{itemize}	
		\item The second is that the data exhibit high reciprocity such that the directed network is effectively a symmetric undirected network. How would the results differ for a more evenly-balanced directed network (i.e. one where the probability of a reciprocal tie is 50\%)? 
		\item To address both issues, the authors could either describe in detail the merits and demerits of this empirical application for comparing AME to existing frameworks, or could draw up a simple simulation to show the sensitivity of the model comparison results to the factors identified above. %Something along the lines of constructing a simulated network with known determinants of tie formation based on nodal characteristics could be helpful here (and could certainly be relegated to the appendix to stay within the word limits). 
	\end{itemize}
\end{enumerate}

\subsection{Minor Comments}

\begin{enumerate}
	\item It would be helpful for the reader to compare AME to non-network models as well, so the authors could include a column in Table 3 for results from a logit model (or the like). 
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have included a logit model for comparison.
		}}
	\end{itemize}		
	\item Typo on page 13, first sentence should read: $y = B'x - |u_i – u_j|$ instead of $y = B'x + |u_i - u_j|$.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have fixed this typo.
		}}
	\end{itemize}		
\end{enumerate}

