\section{Reviewer 2}

\subsection{Major Comments}

\begin{enumerate}
	\item Would it be appropriate to incorporate time dependence in your model given the empirical setup.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for his/her comments. Time dependence can be incorporated into the model via time-dependent regression terms such as lags of the dependent variable or simply time-varying regression parameters. We have clarified this point in the conclusion.
		}}
	\end{itemize}
	\item Talking about inference and using non-experimental data as you do for example on country dyads may be confusing to those who may expect a treatment effects estimator.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have added a clarification to this point in ...
		}}
	\end{itemize}	
	\item It could be useful to compare your model to the current work going on in modelling network interdependence in statistics, engineering and epidemiology, That may make your model more relevant to the broader field.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Since our primary audience are political scientists we wanted to center our discussion around the types of approaches that the field is currently familiar with and AME. However, we very much agree that a significant amount of relevant work is being done in each of these fields, so we have added a footnote to the conclusion pointing to current work and how it relates to the AME.
		}}
	\end{itemize}	
	\item I also recommend including the estimation of the basic logit model as that is still the starting point for many approaching this field.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We have included a Logit model for comparison. Figure 2 in the paper shows that the Logit model performs notable worse in terms of out-of-sample predictive performance than the AME.
		}}
	\end{itemize}	
\end{enumerate}